---
title: "Machine Learning Assignment - Prediction Exercise"
author: "Bolaji Ogundipe"
date: "12/1/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Synopsis
This project work is part of exxercise to fulfill requirement of Machine Learning course 
part of coursera Data Science Specializaton program. It aim to predict the manner in which 
6 participant, with data from accelerometers on the belt, forearm, arm, and dumbell quantifies 
how well of a particular activity they do. This is the "classe" variable in the trainng set.
The prediction model (algorithm) are applied to predict 20 different test cases available in the 
test data and submited in appropriate format to the Coursera Project Prediction Quiz for 
automated grading.

## Background to study
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal
activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take
activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take
geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well
they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6
participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available
from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on
the Weight Lifting Exercise Dataset).

## Exploratory Analysis of Data
The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data is from this source: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har. 

But for the generousity of Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. 
Qualitative Activity Recognition of Weight Lifting Exercises. 
Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.
There would'nt have been this data to perform this analysis. I thank you guys!

## General overview of the data by the author;
Six young health participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different 
fashions: exactly according to the specification (Class A), throwing the elbows to the f(Class B), lifting the dumbbell only halfway 
(Class C), lowering the dumbbell only halfway (Class D) and throwing the hips to the front (Class E).

Class A corresponds to the specified execution of the exercise, while the other 4 classes correspond to common mistakes. 
Participants were supervised by an experienced weight lifter to make sure the execution complied to the manner they were supposed to simulate. 
The exercises were performed by six male participants aged between 20-28 years, with little weight lifting experience. We made sure that all 
participants could easily simulate the mistakes in a safe and controlled manner by using a relatively light dumbbell (1.25kg).

## Setting work environment
```{r, echo=TRUE}
setwd("~/DScoursera/Machine Learning/ML project_work")
set.seed(12345)
library(caret)
library(corrplot)
library(rattle)
library(randomForest)
library(mice)
library(knitr)
library(rpart.plot)
library(rpart)
```
## Loading and Cleaning Data 
```{r, echo=TRUE}
urlTrain <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
urlTest <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

training <- read.csv(url(urlTrain))
test <- read.csv(url(urlTest))

inTrain <- createDataPartition(training$classe, p=0.7, list = FALSE)

train <- training[inTrain, ]
test <- training[-inTrain, ]

dim(train)
dim(test)

#removing almost zero variables
NZV <- nearZeroVar(train)
train <- train[, -NZV]
test <- test[, -NZV]

dim(train)
dim(test)

#removing Na values
NAs <- sapply(train, function(x) mean(is.na(x))) >0.95
train <- train[, NAs==FALSE]
test <- test[, NAs==FALSE]

dim(train)
dim(test)

#removing variables use for identification only
train <- train[, -(1:5)]
test <- test[, -(1:5)]

dim(train)
dim(test)
```
Cleaning procedure reduced the variables to 54. Below is correlation matrix of the varibles. In the graph, dark-color areas shows the highly correlated variables. I would have carried out a PCA to further reduce the number of variables but it is few variables that are highly correlated in this case.
```{r, echo=TRUE}
corMatrix <- cor(train[, -54])
corrplot(corMatrix, order = "FPC", method = "circle", type = "full", tl.cex = 0.8, tl.col = rgb(0, 0, 0))
```
## Prediction Models
*Random Forest*
```{r, echo=TRUE}
set.seed(3245)
controlRF <- trainControl(method="cv", number=3, verboseIter=FALSE)

modfitRF <- train(classe ~., data=train, method = "rf", trControl = controlRF)

modfitRF$finalModel

#prediction on test data
predictRF <- predict(modfitRF, newdata = test)
confmat <- confusionMatrix(predictRF, test$classe)
confmat
```
*Decision Tree*
```{r, echo=TRUE}
set.seed(3245)
modfitDT <- rpart(classe ~., data = train, method = "class")
fancyRpartPlot(modfitDT)
```
predict on test data
```{r, echo=TRUE}
predictDT <- predict(modfitDT, newdata = test, type = "class")
confmatDT <- confusionMatrix(predictDT, test$classe)
confmatDT
```
apply model to data
```{r, echo=TRUE}
predictTest <- predict(modfitRF, newdata = test)
predictTest
```